Batch, Off-Policy Apprencticeship Learning

### 2011-03-28 main.tex
LSTDmu de matthieu

\section{LSTD-$\mu$}
\label{sec:lstdmu}
The similarity between equations \ref{eqn:mu} and \ref{eqn:V} (both are Bellman equations) allows the use of the LSTD algorithm in order to estimate the feature expectation of a policy.\\

We introduce the matrix $\zeta$ such as :
\begin{equation}
\hat \mu_\pi (s) = \zeta^T\phi(s,\pi(s))
\end{equation}
$\zeta$ is given by :
\begin{equation}
\zeta = \tilde A^{-1}\tilde b
\end{equation}
with :
\begin{eqnarray}
\tilde A = \sum_{\{s,a,s'\} \in D}\phi(s,a)\left(\phi(s,a) - \gamma \phi(s',\pi(s'))\right)^T\\
\tilde b = \sum_{\{s,a,s'\} \in D} \phi(s,a)\psi(s)^T
\end{eqnarray}

This approach is superior to the Monte-Carlo estimation suggested by \citet{abbeel2004apprenticeship}, as it allows for a batch, off policy estimation of $\mu_\pi$, even when no model of the environment is known and no simulator is available whereas the Monte-Carlo estimation needs either a model or a simulator, and only allow for an estimation of $\mu_\pi(s_0)$.\\

Coupled with the LSPI algorithm, this allows an adaptation of \citet{abbeel2004apprenticeship} algorithm to "real-life" settings where the only data available come from the expert's trajectories.\\

FIXME: Un mot sur pourquoi il est inutile de faire des itérations sur mu directement ?

\section{Analysis}
FIXME: Matthieu ?
\section{Experimental benchmark}
\label{sec:exp}
The experimental benchmark chosen here is the same as in \citep{ng2000algorithms}, a 5x5 grid world. The agent is in one of the case of the grid (whose coordinates is the state) and can choose at each step one of the four compass directions (the action). With probability 0.9, the agent moves in the intended direction. With probability 0.1, the direction is randomly drawn (and thus have probability 0.25 to match the original direction). The reward optimized by the expert is 0 everywhere but in the upper-right square, where it is 1. For every policy, an episodes ends when the upper right square is attained, or after 20 steps. At the start of each episode, the agent is in the lower-left corner of the grid (the opposite of where the reward is).\\

>>>>

\section{LSTD-$\mu$}
Let us write the definition of the $i^\text{th}$ component of a
feature expectation $\mu^\pi(s)$ for some policy $\pi$:
\begin{equation}
  \mu_i^\pi(s) = E[\sum_{t=0}^\infty \gamma^t \phi_i(s_t)|s_0=s,\pi]
\end{equation}
This is exactly the definition of the value function of the policy
$\pi$ for the MDP considered with the $i^\text{th}$ basis function
$\phi_i(s)$ as the reward function. There exists many algorithms to
estimate a value function, any of them can be used to estimate
$\mu_i^\pi$. Based on this remark, we propose to use specifically
the least-squares temporal difference (LSTD)
algorithm~\citep{bradtke1996linear} to estimate each component of the
feature expectation (as each of these components can be understood
as a value function related to a specific and known reward
function).

Assume that a set of transitions $\{(s_t,r_t,s_{t+1})_{1\leq t \leq
n}\}$ sampled according to the policy $\pi$ is available. We assume
that value functions are searched for in some  hypothesis space
$\mathcal{H}_\psi = \{ \hat{V}_\xi(s) = \sum_{i=1}^q \xi_i \psi_i(s)
= \xi^T \psi(s), \xi\in\mathbb{R}^q\}$. As reward and value
functions are possibly quite different, another hypothesis space is
considered for value function estimation. But if $\mathcal{H}_\phi$
is rich enough, one can still consider
$\mathcal{H}_\psi=\mathcal{H}_\phi$. Therefore, we are looking for
an approximation of the following form:
\begin{equation}
  \hat{\mu}_i^\pi(s) = (\xi_i^*)^T \psi(s)
\end{equation}
The parameter vector $\xi_i^*$ is here the LSTD estimate:
\begin{equation}
  \xi_i^* = \left(\sum_{t=1}^n
  \psi(s_t)(\psi(s_t)-\gamma\psi(s'_{t}))^T\right)^{-1}
  \sum_{t=1}^n \psi(s_t) \phi_i(s_t)
\end{equation}
For apprenticeship learning, we are interested more particularly in
$\hat{\mu}^\pi(s_0)$. Let $\Psi = (\psi_i(s_t))_{t,i}$ be the
$n\times q$ matrix of values predictors, $\Delta\Psi = (\psi_i(s_t)
- \gamma\psi_i(s'_t))_{t,i}$ be the related $n\times q$ matrix and
$\Phi = (\phi_i(s_t))_{t,i}$ the $n\times p$ matrix of reward
predictors. It can be easily checked that $\hat{\mu}^\pi(s_0)$
satisfies:
\begin{equation}
  (\hat{\mu}^\pi(s_0))^T = \psi(s_0)^T (\Psi^T \Delta\Psi)^{-1} \Psi^T \Phi
\end{equation}
This provides an efficient way to estimate the feature expectation
of the expert in $s_0$.

There remains to compute the feature expectations of intermediate
policies, which should be done in an off-policy manner (that is
without explicitly sampling trajectories according to the policy of
interest). To do so, still interpreting each component of the
feature expectation as a value function, we introduce a state-action
feature expectation defined as follows (much as the classic
$Q$-function extends the value function):
\begin{equation}
  \mu^\pi(s,a) = E[\sum_{t=0}^\infty \gamma^t
  \phi(s_t)|s_0=s,a_0=a,\pi]
\end{equation}
Compared to the classic feature expectation, this definition adds a
degree of freedom on the first action to be chosen before following
the policy $\pi$. With a slightly different definition of the
related hypothesis space, each component of this feature expectation
can still be estimated using the LSTD algorithm (namely using the
LSTD-Q algorithm~\citep{lagoudakis2003least}). The clear advantage of
introducing the state-action feature expectation is that this
additional degree of freedom allows off-policy learning. Assuming
that $\mathcal{H}_\psi$ defines state-action feature vectors and
that a set of transitions $\{(s_t,a_t,r_t,s_{t+1})_{1\leq t \leq
n}\}$ sampled according to some behaviorial policy $\pi_0$ is
available, the parameter vector $\xi_i^*$ related to the
$i^\text{th}$ component of the state-action feature expectation
$\mu^\pi(s)$ (policy $\pi$ being different from policy $\pi_0$) is
given by:
\begin{equation}
  \xi_i^* = \left(\sum_{t=1}^n
  \psi(s_t,a_t)(\psi(s_t,a_t)-\gamma\psi(s'_{t},\pi(s'_t)))^T\right)^{-1}
  \sum_{t=1}^n \psi(s_t,a_t) \phi_i(s_t)
\end{equation}
Extending LSTD-$\mu$ to state-action LSTD-$\mu$ is done in the same
manner as LSTD is extended to LSTD-Q~\citep{lagoudakis2003least},
technical details are not provided here for the clarity of
exposition.

Given the (state-action) LSTD-$\mu$ algorithm, the apprenticeship
learning algorithm of~\citep{abbeel2004apprenticeship} (see
Section~\ref{sec:back}) can be easily extended to a batch and
off-policy setting. The solely available data is a set of
transitions sampled according to the expert policy. The
corresponding feature expectation for the starting state $s_0$ is
estimated with the LSTD-$\mu$ algorithm. At step~4 of this
algorithm, the MDP is (approximately) solved using
LSPI~\citep{lagoudakis2003least} (an approximate policy iteration
algorithm using LSTD-Q as the off-policy $Q$-function estimator).
The corresponding feature expectation at state $s_0$ is estimated
using the proposed state-action LSTD-$\mu$.

Before presenting some experimental results, let us stress that
LSTD-$\mu$ is simply the LSTD algorithm applied to a specific reward
function. Although quite clear, the idea of using a temporal
difference algorithm to estimate the feature expectation is new, as
far as we know. A clear advantage of the proposed approach is that
any theoretical result holding for LSTD also holds for LSTD-$\mu$,
such as convergence~\citep{nedic2003least} or
finite-sample~\citep{lazaric2010finiteLSTD} analysis for example.

\section{Experimental benchmark}
\label{sec:exp}
The experimental benchmark chosen here is the same as in \citep{ng2000algorithms}, a 5x5 grid world. The agent is in one of the case of the grid (whose coordinates is the state) and can choose at each step one of the four compass directions (the action). With probability 0.9, the agent moves in the intended direction. With probability 0.1, the direction is randomly drawn (and thus have probability 0.25 to match the original direction). The reward optimized by the expert is 0 everywhere but in the upper-right square, where it is 1. For every policy, an episodes ends when the upper right square is attained, or after 20 steps. At the start of each episode, the agent is in the lower-left corner of the grid (the opposite of where the reward is).\\
